### 2018-07-26 Paper/paper.tex

%&pdflatex
\documentclass[10pt, conference, letterpaper]{IEEEtran}

\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{todonotes}

%\newcommand{\toolname}{BGP-SAS}
%\newcommand{\toolname}{BGP Path Finder}
\newcommand{\toolname}{BGP Path Guide}
\newcommand{\timerange}{(2012-2018)}

\newcommand{\Ben}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommand{\Steph}[2][1=]{\todo[linecolor=yellow,backgroundcolor=yellow!25,bordercolor=yellow,#1]{#2}}

%Titles:
%\title{The Evolution of National AS Chokepoints and Their Connection to
%Internet Freedom}
\title{Borders and Gateways: Measuring and Analyzing National AS Chokepoints}

\graphicspath{{./images/}}

\begin{document}
\maketitle

\input{abstract}
\input{intro}
\input{background}
\input{chokepoint}
\input{tool}
\input{experiment}
\input{results}
\input{related}
\input{conclusion}

\bibliographystyle{plain}
\bibliography{paper}

\end{document}